Sparsity Weak Lensing $3$- Mass Reconstruction

### 2020-08-14 paper_ms_method_HSCY1/ms.tex
add talk

\documentclass[twocolumn]{aastex62}
\emergencystretch=1.em
\newcommand \figPath{./}
\newcommand \redColor{\color{red}}
\newcommand{\vecb}[1]{{#1}}
\renewcommand\labelenumi{(\roman{enumi})}
\renewcommand\theenumi\labelenumi
%\newcommand{\jcap}{JCAP}

\usepackage{amsmath,amsthm,amsfonts,amssymb,bm}
\usepackage{listings}
\usepackage{multirow}
\usepackage{color}

%\usepackage{subfigure}
\usepackage{textcomp}
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{commath}
\hypersetup{breaklinks}
\DeclareMathOperator{\arccosh}{arccosh}

\begin{document}
\title{Sparsity Weak Lensing Mass Reconstruction}
%\author{Xiangchong Li}
%\affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{Masamune Oguri}
%\affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\affiliation{Research Center for the Early Universe, University of Tokyo, Tokyo 113-0033, Japan}
%\author{Nobuhiko Katayama}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{Wentao Luo}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{Wenting Wang}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{Jiaxin Han}
%\affiliation{Department of Astronomy, School of Physics and Astronomy, Shanghai Jiao Tong University, Shanghai, 200240, China}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{HSC Collaboration}
%\noaffiliation
%\email{xiangchong.li@ipmu.jp}

@@

\section{Introduction}
Light from distant galaxies is lensed by the foreground inhomogeneous density distribution along the line-of-sight 
due to the influence of gravity and the shapes of the foreground galaxies are distorted due to the lensing effect. 
Such effect, which is known as weak lensing, imprints the information of foreground density distribution to the 
background galaxies and offers a direct probe into the mass density distribution in the universe.

This paper is organized as follows. 
Section \ref{sec:MethodAll} proposes the new method for $3$-D mass map reconstruction. 
Section \ref{sec:Sim} introduces the realistic simulations we use to test the new method.
Section \ref{sec:Res} presents the results of our method on the simulations.
Section \ref{sec:Sum} summarizes and discusses the future development of the method.

\section{Methodology}
\label{sec:MethodAll}

\subsection{The Problem}
\subsubsection{Density Contrast to Shear}
We first review how the foreground density contrast induces the shear distortion on background galaxies.
The lensing distortion $\kappa$ at a comoving distance $\chi_s$ caused by the foreground inhomogeneous 
density distribution at a comoving distance $\chi_l$ ($\chi_l< \chi_s$) is

\begin{figure*}
    \includegraphics[width=0.95\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The NFW atom with different scale radius ($r_s$). The first row shows the NFW atoms in Fourier 
            space and the second row shows the NFW atom in Real space.}
\end{figure*}

\begin{equation}
\kappa(\vec{\theta},\chi_s)=\frac{3H_0^2\Omega_M}{2 c^2} \int_0^{\chi_s} d\chi_l \frac{\chi_l \chi_{sl}}{\chi_s}
\frac{\delta(\vec{\theta},\chi_l)}{a(\chi_l)},
\end{equation}
\citep{massMap-Glimpse3D2014}.
where $\chi$ refers to the comoving distance, $\delta=\rho(\vec{\theta},\chi_l)/\bar{\rho}-1$ is the density contrast
at the position of lens, $H_0$ is the Hubble parameter, $\Omega_M$ is the matter density parameter, $c$ is the speed 
of light, and $a(\chi_l)$ is the scale parameter at the lens position.

Substitute comoving distance ($\chi$) with redshift ($z$)
$z$ and we have
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l \delta_{c}^{-1}(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
We term $\delta_{c}(z_l,z_s)$ as lensing critical density contrast and define it as
\begin{equation}
\delta_{c}^{-1}(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{\chi_{s} E\left(z_l\right)} & (z_s>z_l),\\
0&(z_s \leq z_l).
\end{cases}
\end{equation}

As shown in \citet{massMap-KS1993}, the shear distortion is related to the kappa field at the redshift plane as
\begin{equation}\label{eq-kappa2gamma}
\vecb{\gamma}(\vec{\theta},z_s) = \int \vecb{D}(\vec{\theta}-\vec{\theta'}) \kappa(\vec{\theta'},z_s) d^2 \theta,
\end{equation}
where
\begin{equation}
\begin{split}
\vecb{D}(\vec{\theta})&=-\frac{1}{\pi}(\theta_1-i\theta_2)^{-2},\\
\vecb{\gamma}(\vec{\theta})&=\gamma_1(\vec{\theta})+i\gamma_2(\vec{\theta}).
\end{split}
\end{equation}

Combine equation (\ref{eq-delta2kappa}) with equation (\ref{eq-kappa2gamma}) and we have
\begin{equation}\label{eq-delta2gamma-z}
\vecb{\gamma}(\vec{\theta},z_s) = \int_0^{z_s} dz_l \frac{\delta_c(z_l,+\infty)}{\delta_{c}(z_l,z_s)} \int d^2 \theta'  \vecb{D}(\vec{\theta}-\vec{\theta'}) \frac{\delta(\vec{\theta'},z_l)}{\delta_c(z_l,+\infty)}.
\end{equation}

\subsubsection{Photo-$z$ Uncertainty}
\begin{figure}
 \includegraphics[width=0.5\textwidth]{nfwlet-atom-1D.pdf}
 \caption{The $1$-D slices of NFW atoms with different scale radius ($r_s=0,1,2,4,8$).}
\end{figure}
Since the redshifts of source galaxies in the current large scale survey are estimated with a limited 
number of photometric bands, the estimated redshift of a single galaxy suffers from large uncertainty. 
We denote the probability of a galaxies, with photo-$z$ estimated as $z_s$, being  located at redshift 
$z$ as $\mathcal{P}(z|z_s)$.  Note that, in order to simplify the future calculation, we assume the 
variation of the PDF across the transverse plane is small and neglect such variation.

Taking the uncertainty of redshift into account, equation (\ref{eq-delta2gamma-z}) changes to
\begin{equation}\label{eq-delta2gamma-poz}
\begin{split}
\gamma(\vec{\theta},z_s) &= \int_0^{z_s} dz_l P(z_l,z_s)\gamma(\vec{\theta},z_l,+\infty),\\
\gamma(\vec{\theta},z_l,+\infty)&= \int d^2 \theta'   Q(\vec{\theta}-\vec{\theta'},z_l) \delta(\vec{\theta'},z_l),
\end{split}
\end{equation}
where $\gamma(\vec{\theta},z_l,+\infty)$ represents shear signal at infinite redshift ($z_s=+\infty$) caused by density contrast at $z_l$, and $P(z_l,z_s)$ is the convolution kernel along line-of-sight and $Q(\vec{\theta},\vec{\theta}',z_l)$ is the convolution kernel along on the transverse plane. These kernels are defined as
\begin{equation}
\begin{split}
P(z_l,z_s)&=\int \frac{\delta_c(z_l,+\infty)}{\delta_{c}(z_l,z)} \mathcal{P}(z|z_s) dz,\\
Q(\vec{\theta},\vec{\theta'},z_l)&=\frac{\vecb{D}(\vec{\theta}-\vec{\theta'})}{\delta_c(z_l,+\infty)}.
\end{split}
\end{equation}

\subsubsection{Smoothing}
Since the observed galaxies have random irregular (unequally-spaced) distribution, it is necessary to smooth the shear signal in the observation. The smoothing is expressed as follows
\begin{equation}
\gamma_S  (\vec{\theta},z)  = \frac{\sum_i  W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) e_i}{\sum_i W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) R_i},
\end{equation}
where $W(\vec{\theta},z)$ is a $3$-D smoothing kernel. $e_i$, $R_i$, $z_i$ and $\theta_i$ are the ellipticity, response, reshift, transverse position of the `$i$-th' galaxy.
$W(\vec{\theta},z)$ can be decomposed into a transverse component $W_T(\vec{\theta})$ and a line-of-sight component $W_\times(z)$
\begin{equation}
W(\vec{\theta},z)=W_T(\vec{\theta}) W_\times (z).
\end{equation}
In this paper, we set
\begin{equation}
\begin{split}
W_T(\vec{\theta}) &=\frac{1}{2\pi\beta^2}\exp(-\frac{|\vec{\theta}|}{2\beta^2}),\\
W_\times (z) &=
\begin{cases}
1/\Delta z& (|z|<\Delta z/2),\\
0& else.
\end{cases}
\end{split}
\end{equation}

With the assumption that the density of response $R$ and the density of galaxy number vary slowly on the smoothing scale, the smoothed shear is
\begin{equation}\label{eq-smooth-gamma}
\gamma_S(\vec{\theta}_j,z_j)= \int d^2 \theta_s d z_sW(\vec{\theta}_j-\vec{{\theta}}_s,z_j-z_s) \gamma(\vec{\theta}_s,z_s)
\end{equation}

Substitute equation (\ref{eq-delta2gamma-poz}) into equation (\ref{eq-smooth-gamma})
\begin{equation}\label{eq-delta2gamma-smooth}
\begin{split}
\gamma_S(\vec{\theta}_j,z_j) &= \int_0^{z_j} dz_l P_S(z_l,z_j)\gamma_S(\vec{\theta}_j,z_l,+\infty),\\
\gamma_S(\vec{\theta}_j,z_l,+\infty)&= \int d^2 \theta'   Q_S(\vec{\theta}_j,\vec{\theta'},z_l) \delta(\vec{\theta'},z_l),
\end{split}
\end{equation}
and
\begin{equation}
\begin{split}
P_S(z_l,z_j)&=\int  dz_s W_\times (z_j-z_s) \int \frac{\delta_c(z_l,+\infty)}{\delta_{c}(z_l,z)} \mathcal{P}(z|z_s) dz,\\
Q_S(\vec{\theta},\vec{\theta'},z_l)&=\int d^2\theta'' W_T( \vec{\theta} -\vec{\theta''} )  \frac{\vecb{D}(\vec{\theta''}-\vec{\theta'})}{\delta_c(z_l,+\infty)}.
\end{split}
\end{equation}

\subsubsection{Mask and Noise}
In real observations, the influence of mask and noise should also be taken into account. The final observed shear is
\begin{equation}\label{eq-delta2gamma-final}
\gamma_o(\vec{\theta},z)=M(\vec{\theta},z) \gamma(\vec{\theta},z) + \epsilon(\vec{\theta},z),
\end{equation}
where $\epsilon$ represents noise typically caused by random shape of intrinsic galaxies and such noise is modeled as white Gaussian noise. $M(\vec{\theta},z_s)$ is the mask function.

We can write equation (\ref{eq-delta2gamma-final}) into
\begin{equation}\label{eq-delta2gamma-operator-final}
\gamma_o =\mathbf{M}  \mathbf{P} \mathbf{Q} \delta + \epsilon,
\end{equation}
where $\mathbf{M}$,  $\mathbf{P}$ and $\mathbf{Q}$ represent the functional operators $M$, $P$ and $Q$, respectively.

\subsection{The Method}
\label{sec:kappaMap}

\subsubsection{Dictionary}
Since many $N$-body simulations have shown the dark matter to be largely distributed in halos connected by filaments,
we assume that the density contrast field can be decomposed into multi-scaled NFW halo \citep{halo-NFW1997ApJ} and point mass at different positions
\begin{equation}
\delta(\vec{x}) = \sum_{s=0}^{N_s} \int d^3 x' \phi_s(\vec{x}-\vec{x}') \alpha_s(\vec{x'})\\
\end{equation}
where $\phi_0$ is $3$-D Dirac delta function representing point mass
\begin{equation}
\phi_0(\vec{x})= \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}
Since the scale of halo is much less than the reachable redshift resolution, we neglect the depth of halo on the line-of-sight direction as suggested by \citep{massMap-Glimpse3D2014}.
On the transverse plane, the NFW atom is modeled with the surface density profile of NFW halo with scale $\theta_s$ and truncation radius $c \theta_s$ \citep{haloModel-TJ2003-3pt}, where $c$ is generally known as concentration of NFW halo.  Whereas, the NFW atom is modeled with Dirac delta function on the line-of-sight direction

\begin{equation}
\begin{split}
\phi_s(\vec{x}) =&\frac{f }{2 \pi \theta_s^2 } F(|\vec{\theta}|/\theta_s) \delta_D(z),\\
&  (s=1..N_s)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh \left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
and $f=1/[\ln (1+c)-c/(1+c)]$.

$\alpha_s$ is the corresponding decomposition coefficient of the density contrast field onto our basis atom. The total coefficients set is denoted as $\alpha=\begin{pmatrix}
\alpha_0\\
\alpha_1\\
...\\
\alpha_{N_s}
\end{pmatrix}$, and the total dictionary is $\Phi=\begin{pmatrix}
\phi_0, \phi_1, ..., \phi_{N_s}
\end{pmatrix}$.

We pixelize the parameter space into a $N_\theta \times N_\theta \times N_l$ grid. $N_\theta$ is the number of pixels on the two dimensions of the transverse plane and $N_l$ is the number of pixels on the line-of-sight direction for the lenses. Similarly, $\gamma_o$ is pixelized onto a $N_\theta \times N_\theta \times N_s$ grid, where $N_s$ is the number of pixel on the line-of-sight direction for the sources.
The pixel size for the two dimensions on the transverse plane is denoted as $\Delta \theta$ and the pixel size for the line-of-sight direction is denoted as $\Delta z$.
Equation (\ref{eq-delta2gamma-operator-final}) changes to
\begin{equation}\label{eq-alpha2gamma-operator}
\gamma=\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} \alpha +\epsilon.
\end{equation}

\subsubsection{Loss Function with Constrains}
\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their mlz best photo-$z$ estimation. The blue histogram is the number distribution of the best photo-$z$ estimation. The vertical dashed lines are the bounds of bins. The galaxies are evenly distributed in each bins.}
\end{figure}

The loss function is defined as
\begin{equation}\label{eq-lossFun}
L = \norm{\gamma-\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} \alpha}^2_2+\lambda \norm{\alpha}^1_1 + \tau\mathrm{TSV}(\alpha_0),
\end{equation}
where $\norm{\gamma-\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} \alpha}^2_2$ is the normal chi-square, $\norm{\alpha^{N,P}}^1_1$ is the sparsity constrain, and $\mathrm{TSV}(\alpha^P)$ is the Total Squared Variation (TSV) constrain. $\norm{\bullet}_1$ and $\norm{\bullet}_2$ represent $l_1$ norm and $l_2$ norm, respectively.
The $\mathrm{TSV}$ operator is defined as
\begin{equation}
\begin{split}
\mathrm{TSV}(\alpha_0) =&  \sum_{ijk} \{ (\alpha_0[i,j,k]-\alpha_0[i+1,j,k])^2\\
&+(\alpha_0[i,j,k]-\alpha_0[i,j+1,k])^2\\
&+(\alpha_0[i,j,k]-\alpha_0[i,j,k+1])^2 \},
\end{split}
\end{equation}
where $i=1..N_\theta$, $j=1..N_\theta$, are the pixel indexes for the two dimensions on the transverse plane, $k=1..N_l$ is the pixel index on the line-of-sight direction.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{mlz-poz.pdf}
 \caption{The PDF of photo-$z$ error for $10$ source redshift bins.}
\end{figure}

>>>>

\documentclass[twocolumn]{aastex62}
\emergencystretch=1.em

\usepackage{amsmath,amsthm,amsfonts,amssymb,bm}
% for big dot
\makeatletter
\newcommand*\bigcdot{\mathpalette\bigcdot@{.5}}
\newcommand*\bigcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}

\usepackage{listings}
\usepackage{multirow}
\usepackage{color}

%\usepackage{subfigure}
\usepackage{textcomp}
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{commath}
\hypersetup{breaklinks}
\DeclareMathOperator{\arccosh}{arccosh}
\newcommand \figPath{./}
\newcommand \redColor{\color{red}}
\renewcommand\labelenumi{(\roman{enumi})}
\renewcommand\theenumi\labelenumi

\usepackage{algorithm,algcompatible}
\DeclareMathOperator*{\argmax}{\arg\!\max}
\algnewcommand\INPUT{\item[\textbf{Input:}]}
\algnewcommand\OUTPUT{\item[\textbf{Output:}]}

\makeatother
%\newcommand{\jcap}{JCAP}

\begin{document}
\title{Sparsity Weak Lensing $3$- Mass Reconstruction}
%\author{Xiangchong Li}
%\affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{Masamune Oguri}
%\affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\affiliation{Research Center for the Early Universe, University of Tokyo, Tokyo 113-0033, Japan}
%\author{Wentao Luo}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{HSC Collaboration}
%\noaffiliation
%\email{xiangchong.li@ipmu.jp}

@@

\section{Introduction}
Light from distant galaxies is distorted by the intervening inhomogeneous density distribution along the line-of-sight
due to the influence of gravity and the shapes of the background galaxies are coherently sheared due to the lensing effect.
Such effect, which is known as weak lensing, imprints the information of foreground mass density distribution to the
background galaxy images and offers a direct probe into the mass density distribution in the universe \citep[see][for recent 
reviews]{revKilbinger15,revRachel17}.

Mathematically, the shear field ($\gamma$) measured distant galaxies is a linear mapping of the density contrast ($\delta$)
\begin{equation}
 \gamma=\mathbf{T} \delta,
\end{equation}
where $\mathbf{T}$ is the transformation operator. The transformation operator ($\mathbf{T}$) includes not only physical lensing
effect but it also includes observational systematic effects.

In order to fully reconstruct the $3$-D mass density distribution ($\delta$) from the $3$-D shear field ($\gamma$) 
observed from distorted galaxy images, a premise is take that the density field can be decomposed into a set of models
\begin{equation}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the model space to the mass density and $x$ is the projector. 
In the previous works, Starlets (wavelets transform) and sinusoidial functions (Fourier transform) have been used to 
modeled the density field .

Subsequently, an ill-posed inverse problem
\begin{equation}
\rm{min}_x\{\norm{\gamma - \mathbf{T\Phi} x}_2^2+ C(x)\},
\end{equation}
need to be solved to reconstruct the mass density map, where $C(x)$ is the regulation term which makes the ill-posed 
inversion solvable. The $l^1$ sparsity regulation and the $l^2$ ridge regulation have been applied to the mass density 
map reversion problem. In their application, the $l^1$ sparsity regulation ($\norm{x}^1_1$) is combined with Starlets modeling 
\citep{LSS-massMap-Glimpse3D-Leonard2014} and the $l^2$ ridge regulation ($\norm{x}^2_2$) is combined with sinusoidial modeling\footnote{It 
is also known as Wiener filter.} \citep{LSS-massMap-Wiener-Simon2009}. See \citet{HSC1-massMaps} for a recent application 
of \citet{LSS-massMap-Wiener-Simon2009} in the Hyper Suprime-Cam Survey \citep{HSC1-data}.

We propose to use NFW halos \citep{halo-NFW1997ApJ} to model the density contrast field and construct the density contrast 
field with $l^1$ sparsity regulation.

This paper is organized as follows.
Section \ref{sec:Method} proposes the new method for $3$-D mass map reconstruction.
Section \ref{sec:Sim} introduces the realistic simulations we use to test the new method.
Section \ref{sec:Res} presents the results of our method on the simulations.
Section \ref{sec:Sum} summarizes and discusses the future development of the method.

\section{Methodology}
\label{sec:Method}

In this section, we first review how the foreground density contrast induces the weak lensing shear distortion on background 
galaxies in section \ref{subsec:method-delta2shear}.
After that, we describe several systematic effects which exist in real observation mathematically. The systematic effects 
include photo-$z$ uncertainty (section \ref{subsec:method-photoz}), smoothing (section \ref{subsec:method-smoothing}),
mask and noise (section \ref{subsec:method-msknoise}) and pixelation (section \ref{subsec:method-pix}).

Subsequently, we propose a novel method to reconstruct the $3$-D density contrast field from the weak lensing shear 
distortion field. 
With the premise that the density field can be decomposed into NFW halos and point masses, we build up a dictionary using 
NFW models and point mass models in section \ref{subsec:method-dictionary}.
The models in our dictionary are used to fit the shear field and the optimal fit is determined by minimizing a loss function. 
As shown in section \ref{subsec:method-lossfun} the loss function is composed of a normal chi-$2$ term, a $l_1$ sparsity 
constrain and a quadratic total square variance constrain.
The minimum of the loss function is achieved with the pathwise coordinate descent algorithm in section \ref{subsec:method-pathwise}.

\subsection{Density Contrast to Shear}
\label{subsec:method-delta2shear}

\begin{figure*}
    \includegraphics[width=0.95\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The NFW models with different scale radius ($r_s$). The first row shows the NFW models in Fourier
            space and the second row shows the NFW model in Real space.}
\end{figure*}

The lensing convergence $\kappa$ at the comoving distance $\chi_s$ caused by the foreground inhomogeneous
density distribution at the comoving distance $\chi_l$ ($\chi_l< \chi_s$) along the line-of-sight is
\begin{equation}
\kappa(\vec{\theta},\chi_s)=\frac{3H_0^2\Omega_M}{2 c^2} \int_0^{\chi_s} d\chi_l \frac{\chi_l \chi_{sl}}{\chi_s}
\frac{\delta(\vec{\theta},\chi_l)}{a(\chi_l)},
\end{equation}
\citep{LSS-massMap-Glimpse3D-Leonard2014}.
where $\chi$ refers to the comoving distance, $\delta=\rho(\vec{\theta},\chi_l)/\bar{\rho}-1$ is the density contrast
at the position of lens, $H_0$ is the Hubble parameter, $\Omega_M$ is the matter density parameter, $c$ is the speed
of light, and $a(\chi_l)$ is the scale parameter at the lens position.

Substitute comoving distance ($\chi$) with redshift ($z$)
$z$ and we have
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l \delta_{c}^{-1}(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
$K(z_l,z_s)$ as the lensing kernel defined as
\begin{equation}
K(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{\chi_{s} E\left(z_l\right)} & (z_s>z_l),\\
0&(z_s \leq z_l).
\end{cases}
\end{equation}

As shown in \citet{massMap-KS1993}, the shear distortion is related to the kappa field at the redshift plane as
\begin{equation}\label{eq-kappa2gamma}
\gamma^t(\vec{\theta},z_s) = \int  d^2 \theta' D(\vec{\theta}-\vec{\theta'}) \kappa(\vec{\theta'},z_s),
\end{equation}
where
\begin{equation}
D(\vec{\theta})=-\frac{1}{\pi}(\theta_1-i\theta_2)^{-2}.
\end{equation}
Here we denote the true shear distortion as $\gamma^t$ to distinguish the observed shear signal with systematics which 
will be introduced in the following subsections.

Combining equation (\ref{eq-delta2kappa}) with equation (\ref{eq-kappa2gamma}), the true shear signal is derived as
\begin{equation}\label{eq-delta2gamma-z}
\gamma^t(\vec{\theta},z_s) = \int_0^{z_s} \frac{dz_l}{\delta_{c}(z_l,z_s)} \int d^2 \theta' \vec{D}(\vec{\theta}-\vec{\theta'}) \delta(\vec{\theta'},z_l).
\end{equation}
Here we define the lensing functional as 
\begin{equation}
\mathbf{Q}=\int_0^{z_s} \frac{dz_l}{\delta_{c}(z_l,z_s)} \int d^2 \theta'  \vec{D}(\vec{\theta}-\vec{\theta'}),
\end{equation}
then eq. (\ref{eq-delta2gamma-z}) is simplified to $\gamma^t=\mathbf{Q}\delta$.

\subsection{Photo-$z$ Uncertainty}
\label{subsec:method-photoz}
\begin{figure}
 \includegraphics[width=0.5\textwidth]{nfwlet-atom-1D.pdf}
 \caption{The $1$-D slices of NFW models with different scale radius ($r_s=0,1,2,4,8$).}
\end{figure}

Since the photometric redshifts of source galaxies in the current large scale survey are estimated with 
a limited number of photometric bands, the estimated redshift of a single galaxy suffers from large uncertainty.
We denote the probability of a galaxies, with photo-$z$ estimated as $z_s$, being actually located at redshift
$z$ as $P(z|z_s)$.  Note that, in order to simplify the future calculation, we assume the variation of the PDF 
across the transverse plane is small and neglect such variation.

Taking the uncertainty of redshift into account, the shear signal changes to
\begin{equation}\label{eq-delta2gamma-poz}
\gamma^p(\vec{\theta},z) = \int dz_s P(z|z_s) \gamma^t(\vec{\theta},z_s).
\end{equation}
With the definition of photo-$z$ functional
\begin{equation}
\mathbf{P} = \int dz_s P(z|z_s),
\end{equation}
the relation between shear signal and density contrast is $\gamma^p=\mathbf{P} \mathbf{Q} \delta$.

\subsection{Smoothing}
\label{subsec:method-smoothing}

Since the observed galaxies have random irregular (unequally-spaced) distribution, it is necessary to smooth 
the shear signal in the observation. The smoothing is expressed as follows
\begin{equation}
\hat{\gamma}  = \frac{\sum_i  W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) e_i}{\sum_i R_i W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) },
\end{equation}
where $W(\vec{\theta},z)$ is a $3$-D smoothing kernel. $e_i$, $R_i$, $z_i$ and $\theta_i$ are the ellipticity, 
response, reshift, and transverse position of the `$i$-th' galaxy in the galaxy catalog.

$W(\vec{\theta},z)$ can be decomposed into a transverse component $W_T(\vec{\theta})$ and a line-of-sight component 
$W_\times(z)$
\begin{equation}
W(\vec{\theta},z)=W_T(\vec{\theta}) W_\times (z).
\end{equation}
In this paper, we set
\begin{equation}
\begin{split}
W_T(\vec{\theta}) &=\frac{1}{2\pi\beta^2}\exp(-\frac{|\vec{\theta}|}{2\beta^2}),\\
W_\times (z) &=
\begin{cases}
1/\Delta z& (|z|<\Delta z/2),\\
0& else.
\end{cases}
\end{split}
\end{equation}

With the approximation that the density of response $R$ and the density of galaxy number vary slowly on the smoothing scale
and the fact that $\int d^3r W(\vec{r})=1$, the smoothed shear is
\begin{equation}\label{eq-smooth-gamma}
\gamma^s(\vec{r})= \int d^3 r' W(\vec{r}-\vec{r'}) \gamma^p(\vec{r'})
\end{equation}

We define the smoothing functional as
\begin{equation}
\mathbf{W} = \int d^3 r' W(\vec{r}-\vec{r'}),
\end{equation}
and the relation between shear signal and density contrast is $\gamma^s=\mathbf{W} \mathbf{P} \mathbf{Q} \delta$.

\subsection{Mask and Noise}
\label{subsec:method-msknoise}

In real observations, the influence of mask and noise should also be taken into account. The final observed shear is
\begin{equation}\label{eq-delta2gamma-final}
\gamma(\vec{\theta},z)=M(\vec{\theta},z) \gamma^s(\vec{\theta},z) + \epsilon(\vec{\theta},z),
\end{equation}
where $\epsilon$ represents noise typically caused by random shape of intrinsic galaxies and such noise is modeled 
as uncorrelated Gaussian noise. $M(\vec{\theta},z_s)$ is the mask function.

With the definition of masking functional,
\begin{equation}
\mathbf{M}= \int d^3 r' M(\vec{r'}) \delta_D(\vec{r}-\vec{r'}),
\end{equation}
where $\delta_D(\vec{r})$ is $3$-D Dirac delta function, we can write equation (\ref{eq-delta2gamma-final}) into
\begin{equation}\label{eq-delta2gamma-operator-final}
\gamma =\mathbf{M} \mathbf{W} \mathbf{P} \mathbf{Q} \delta + \epsilon.
\end{equation}

\subsection{Dictionary}
\label{subsec:method-dictionary}
\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their mlz best photo-$z$ estimation. 
        The blue histogram is the number distribution of the best photo-$z$ estimation. The vertical dashed lines 
        are the bounds of bins. The galaxies are evenly distributed in each bins.}
\end{figure}

$N$-body simulations show that the dark matter to be largely distributed in halos connected by filaments, therefore,
we assume that the density contrast field can be decomposed into multi-scaled NFW halo \citep{halo-NFW1997ApJ} and
point mass at different positions
\begin{equation}
\delta(\vec{r}) = \sum_{s=0}^{N} \int d^3 r' \phi_s(\vec{r}-\vec{r'}) x_s(\vec{r'})\\
\end{equation}
where $\phi_0$ is $3$-D Dirac delta function to model point masses
\begin{equation}
\phi_0(\vec{r})= \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}
Since the scale of halo is much less than the reachable redshift resolution, we neglect the depth of halo on the 
line-of-sight direction following \citep{LSS-massMap-Glimpse3D-Leonard2014}. On the other hand, in the transverse directions, 
the surface density profiles of NFW halos with scale $\theta_\alpha$ and truncation radius $c \theta_\alpha$ 
\citep{haloModel-TJ2003-3pt} are used to model halos, where $c$ is generally known as concentration of NFW halo.

\begin{equation}
\begin{split}
\phi_\alpha(\vec{r}) =&\frac{f }{2 \pi \theta_\alpha^2 } F(|\vec{\theta}|/\theta_\alpha) \delta_D(z),\\
&  (s=1..N)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh \left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$, and $x_\alpha$ is the corresponding projection coefficient of the density contrast
field onto our basis NFW model.

We define the projection parameters set as $x=\begin{pmatrix}
x_{[0]}\\
x_{[1]}\\
...\\
x_{[N]}
\end{pmatrix}$, and the dictionary transforming functional as
\begin{equation}
\mathbf{\Phi}=\begin{pmatrix}
\int d^3r\phi_0(\vec{r}) ~\int d^3r \phi_1(\vec{r})~ ...~\int d^3r \phi_{N}(\vec{r})
\end{pmatrix}.
\end{equation}
Then we write equation (\ref{eq-delta2gamma-operator-final}) into
\begin{equation}\label{eq-alpha2gamma-operator}
\gamma=\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} x +\epsilon,
\end{equation}
where $\epsilon(\vec{r})$ is the error from the shear measurement which includes contributions from both random
galaxy orientations and photon noise.

For simplicity, we denote $\mathbf{A}=\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} $ and eq. (\ref{eq-alpha2gamma-operator})
reduces to 
\begin{equation}\label{eq-x2gamma-simple}
\gamma=\mathbf{A} x +\epsilon.
\end{equation}

\subsection{Pixelization}
\label{subsec:method-pix}
We pixelize the smoothed shear field ($\gamma$) into a $N_\theta \times N_\theta \times N_s$ grid, where $N_\theta$ 
is the number of pixels of two orthogonal axes of the transverse plane and $N_s$ is the number of pixels of the 
line-of-sight axis. $\gamma_{\alpha}$ is used to denote the value recorded on pixel with index $\alpha$, where $\alpha=
1..N_\theta \times N_\theta \times N_s$. The grid on the transverse plan is equally spaced so that we can perform the Fast 
Fourier Transform (FFT) on the transverse plan.
